### 2022-08-03 main.tex
bib and title

\usepackage{natbib}
\bibliographystyle{unsrt}
\date{\today}

@@

\begin{flushleft}{
	\Large
	\textbf\newline{
		Intermediate levels of asymptomatic transmission lead to the highest levels of fatalities \swp{89/96 characters}
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1},
Jonathan Dushoff\textsuperscript{2,3,4},
Bryan T. Grenfell\textsuperscript{1,5},
Joshua S. Weitz\textsuperscript{6,7,8,*}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Princeton School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\textbf{6} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{7} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{8} Institut de Biologie, \'{E}cole Normale Sup\'{e}rieure, Paris, France
\\
\bigskip

@@

Asymptomatic transmission and fatalities \swp{40/40 characters}

@@

Asymptomatic infections improve individual-level outcomes but increase overall transmission, thereby causing more deaths. \swp{121/125 characters}

@@

Asymptomatic infections have hampered the ability to characterize and prevent the transmission of SARS-CoV-2 throughout the pandemic.
Although asymptomatic infections reduce severity at the individual level, they can make population-level outcomes worse if asymptomatic individuals---unaware they are infected---transmit more than symptomatic individuals.
Using an epidemic model, we show that intermediate levels of asymptomatic infection lead to the highest levels of epidemic fatalities when the increase in asymptomatic transmission, due either to individual behavior or mitigation efforts, is strong. 
We generalize this result to include presymptomatic transmission, showing how intermediate levels of non-symptomatic transmission lead to the highest levels of fatalities.
Finally, we extend our framework to illustrate how the intersection of asymptomatic spread and immunity profiles determine epidemic trajectories, including population-level severity, of future variants.  \swp{125/125 words}

@@

SARS-CoV-2 has had devastating effects at the population level.
However, many individuals experienced mild cases, making it harder to estimate the magnitude of spread and fatality rate \citep{nogrady2020data}.
The ratio of fatalities to documented cases (the case-fatality rate, CFR) is typically between 1\%--4\%, varying across population because of testing patterns, treatment practice, case definitions, and other factors \citep{rajgor2020many,VERITY2020669,yang2020early}.
But many infections are never documented;
the ratio of fatalities to total infections (the infection fatality rate, IFR) has been estimated to be closer to 0.5\%--1\% for pre-vaccinated populations whose demographics are similar to those of the United States \citep{levin2020assessing}. 
This means that more than 99\% of individuals infected with COVID-19 will survive. 
Moreover, at least half of the infections are sufficiently mild that they could be classified as subclinical or even asymptomatic.

@@

Early in the pandemic, a COVID-19 outbreak on the Diamond Princess cruise ship played a critical role in understanding the role of asymptomatic infections in the spread of SARS-CoV-2;
the outbreak occurred among 3711 passengers and crew, of whom 634 individuals tested positive by 20 February 2020 \citep{mizumoto2020estimating}.
It has been estimated that 75\% (95\% C: 70\%-78\%) of all infections on the cruise ship were asymptomatic (\fref{evidence}A) with about half of total infections undetected \citep{emery2020}.
The relative transmission rate of asymptomatic individuals aboard the Diamond Princess was not well constrained, but low relative transmission rate (below 25\%) by asymptomatic individuals was ruled out because it required unrealistically high transmissibility for symptomatic individuals (\fref{evidence}B).
Modeling studies have typically assumed that transmissibility is lower for asymptomatic than for symptomatic individuals; assumptions have ranged from 10\%--100\% \citep{ferretti2020quantifying,lavezzo2020}.
Similarities in viral load trajectories of asymptomatic and symptomatic individuals provide indirect support for the transmissibility of asymptomatic individuals (\fref{evidence}C, \citep{Kissler2020}); 
however, differences between inferred total viral load from Ct values and infectious viral load add uncertainties to how well asymptomatic individuals can transmit relative to that of symptomatic individuals \citep{romero2021ct}.
We note also that asymptomaticity is expected to be more heterogeneous in a diversity of outbreak settings \citep{alene2021magnitude}.
For example, during the early pandemic, Davies \textit{et al.}'s analyses of surveillance data across six countries revealed that older individuals were less likely to have subclinical infections (\fref{evidence}D), providing indirect evidence for heterogeneity in asymptomaticity \cite{davies2020}.
Differences in contact rates between age classes further contribute to the heterogeneity in asymptomatic transmissibility.
For now, we primarily focus on a homogeneous population and return to the age effect in discussing our model-based findings.

Despite quantitative uncertainties in asymptomatic transmissibility, individuals infected asymptomatically with SARS-CoV-2 can still transmit to others. 
This means that the presence of asymptomatic infections may have countervailing effects at the population level. 
On one hand, an asymptomatic infection means that the individual infected avoids hospitalization and death. 
On the other hand, asymptomatic infections are less likely to be detected \citep{fraser2004factors,moghadas2020implications}, meaning that asymptomatic individuals are less likely to take precautions and relatively more likely to infect others;
asymptomatic SARS-CoV-2 infections present additional challenges to managing overall disease burden due to the possibility of long COVID \citep{xie2022long}.
Altogether, the prevalence of asymptomatic infections could paradoxically make population-level outcomes worse than if SARS-CoV-2 was more dangerous at the individual level.

@@

High values of $\delta$ required for the nonlinear effects of asymptomaticity on deaths may seem unrealistic.
For this particular model, it does not make biological sense for $\delta$ to be greater than the amount of post-symptomatic transmission, because pre-symptomatic transmission is implicitly included in the $I_s$ compartment.
While several studies have estimated the proportion of pre-symptomatic transmission to be around 30\%--60\% for the SARS-CoV-2 wildtype strain, many of these were likely affected by intervention and behavioral effects, as they were conducted after SARS-CoV-2 awareness became widespread \citep{he2020temporal}.
Instead, \cite{sender2021unmitigated} recently estimated that the proportion of pre-symptomatic transmission could have been as low as 20\% (95\%CI: 6\%--32\%) during the first few weeks of the pandemic when the pandemic-awareness and intervention measures were minimal.
There are two implications of this updated estimate---first, a low proportion of pre-symptomatic transmission suggests that high $\delta$ values are feasible (although not necessarily likely) during the initial pandemic phase; and second, intermediate levels of behavioral effects ($\delta > 0$) would have been already present early in the pandemic to reduce the proportion of pre-symptomatic transmission from 80\% to as low as 40\%.

@@

We now apply our framework to understand the impact of immunity on total fatalities at the population scale by dividing the population into two groups: immunologically naive and protected.
For simplicity, we do not distinguish whether the immunity is derived from natural infections or vaccines.
The dynamics of immunologically naive individuals are equivalent to our original model (\fref{base}).
The dynamics of protected individuals include three additional parameters, which characterize the amount of protection against infection $\epsilon_i$, symptoms (given infection) $\epsilon_s$, and deaths (given symptoms) $\epsilon_d$ (\fref{immune}).
For simplicity, we assume that the population is split in half (50\% naive and 50\% protected) and mixes homogeneously. We also do not consider the separate effect of immunity on transmission (beyond the effect on infection). 
In other words, we assume that asymptomatic infections in protected and unprotected people have the same reproduction numbers (and likewise for the symptomatic infections).
In practice, both asymptomatic and symptomatic infections in protected people are less likely to transmit than their unprotected counterparts \citep{lipsitch2020understanding}: asymptomatic infections in protected people may indicate limited viral replication or even immune boosting, in which case an exposed individual may successfully fight off the pathogen early in infection before it can be transmitted; and symptomatic infections in protected people may reflect a strong immune response (rather than high viral load), in which case symptomaticity can be a poor proxy for transmission.
We assume a relatively strong behavioral effect $\delta=0.8$ for illustration.

@@

The outcomes in our simulations of invading variants resemble the dynamics of the SARS-CoV-2 Omicron variant.
Despite moderate levels of vaccine effectiveness against symptomatic and reduced levels of severe cases caused by the Omicron variant, especially after booster shots \citep{andres2022omicron}, both vaccine- and infection-derived immunity provided limited protection against infections \citep{pearson2021omicron}.
This immune evasion helped the Omicron variant to cause more infections in South Africa than previous variants \citep{sun2022omicron}.
Moreover, even though the Omicron variant is probably milder than the Delta variant \citep{MENNI20221618,ulloa2022estimates}, the number of hospitalizations and deaths caused by the Omicron variant was higher than those caused by the Delta variant in many locations \citep{Iacobuccio254,faust2022omicron,sigal2022estimating}.

There are several limitations to our analysis.
First, while we are able to generalize the model to include both pre-symptomatic and asymptomatic transmission, behavioral and intervention effects must be relatively large in order for the fatality to peak at intermediate levels of asymptomaticity (typically requiring a reduction in transmission rate of 60\% or more for most of our chosen parameter sets).
Second, the model framework is able to incorporate the impacts of immunity of infection, symptoms, and severity,
but we neglected the effects of immunity on transmission, which also has important effects on disease dynamics \citep{saad2020immune,gallagher2021indirect}.
In particular, if immunity provides stronger protection against transmission among immune individuals, population-level outcomes will be better than what our model predicts.
Estimating protection against different endpoints (e.g., infection, symptom, death, and transmission) can help narrow this uncertainty.
Finally, we assumed that asymptomatic and symptomatic individuals are infected for the same amount of time.
Analysis of viral load trajectories suggests that asymptomatic individuals may clear infections faster \citep{Kissler2020};
however, asymptomatic individuals may still transmit for a longer period of time if symptomatic individuals self-isolate quickly after symptom onset.
The individual-level differences in the asymptomatic and symptomatic transmission time scale can have important implications for the inferences and predictions of pathogen dynamics \citep{park2020time,harris2022time};
nonetheless, we expect that predictions on the final size of the epidemic and total fatalities will be robust to small differences in the transmission time scale between asymptomatic and symptomatic individuals.

Even though we assumed a homogeneous population throughout, our analysis also has important implications for age-dependent heterogeneity in asymptomaticity (as shown in \fref{evidence}D).
For example, vaccinations and intervention measures primarily targeting older individuals can prevent severe infections and improve individual-level outcomes.
However, asymptomatic individuals, especially younger individuals with high contact rates, can still transmit to other, older individuals, potentially making population-level outcomes worse than they would be if intervention measures were distributed differently.
We note that other factors, such as the efficacy of a vaccine and types of immunity provided by the vaccine, also play critical roles in making these decisions---in many cases, protecting the most vulnerable will be the optimal decision to minimize deaths \citep{moore2021modelling}.

In summary, using a series of simplified models, we have shown that asymptomatic infections (or, more generally, non-symptomatic transmission) can represent a double-edged sword leading to a better outcome for many individuals while facilitating onward transmission that leads to a worse outcome for the population as a whole.
Extending our framework further shows that immunity profile (i.e., reduction of infection, symptoms, and/or severity due to immunity) plays a critical role in determining the dynamics of future variants;
these results extend previous work on post-pandemic trajectories that focus primarily on cross-immunity \citep{kissler2020projecting,lavine2021immunological}.
For example, while protection against symptoms unaccompanied by protection against transmission protects health at the individual level, it can lead to more infections, and potentially more deaths, at the population level.
A similar concern was raised in prioritizing vaccine choices that could reduce severe outcomes vs. others that could reduce transmission \citep{koelle2022changing}.

SARS-CoV-2 has proven hard to control in large part because transmission is often decoupled from symptoms. 
% Mitigation efforts have often prioritized responding to symptoms---including symptom-based testing, fever checks, mask-wearing for infectious individuals---a different approach that strives to reduce the chance of asymptomatic transmission while increasing treatment of symptomatically infected individuals could both reduce infection risk at the source and the overall rates of severe outcomes for the population as a whole.
Our model reinforces the need for dual approaches---prioritizing the reduction of asymptomatic spread (e.g., via risk awareness campaigns \citep{chande2020real,sinclair2021pairing}, asymptomatic testing programs \citep{mina2021covid,gibson2022surv}, mask-wearing indoors and in crowded environments \citep{jones2020two,prather2020reducing,howard2021ev}, and through improvements in ventilation \citep{allen2021indoor,wang2021airborne}) while improving the treatment of symptomatic cases, particularly amongst older individuals at highest risk for severe outcomes.
Given the link between age and asymptomatic infections \citep{davies2020}, interventions may consider different approaches in strongly age-structured populations (e.g., schools or long-term care facilities).
Mass vaccination is also expected to be important especially if future vaccines induce more transmission blocking.
As more variants continue to emerge, monitoring the impacts of preexisting immunity (whether through vaccination and/or infections  \citep{lopman2021framework}) on preventing infections, and not just disease, will be critical to controlling the course of the pandemic.

@@

First, we consider a compartmental model with asymptomatic and symptomatic infections in a homogeneously mixing population.
The model dynamics are as follows:
\begin{align}
\dot{S} &= -\beta_a S I_a -(1-\delta) \beta_s S I_s \\
\dot{E} &= \beta_a S I_a + (1-\delta) \beta_s S I_s - \nu E\\
\dot{I}_a &= p \nu E - \gamma_a I_a\\
\dot{I}_s &= (1-p) \nu E -\gamma_s I_s\\
\dot{R} &= \gamma_a I_a + (1-f) \gamma_s I_s \\
\dot{D} &= f \gamma_s I_s
\end{align}
where the transmission rate $\beta$ and removal rate $\gamma$ can be potentially differ between asymptomatic and symptomatic individuals.  
Here, $\delta$ denotes the reduction in transmissibility due to responsive measures taken by symptomatic individuals.
Throughout the paper, we use parameters that are broadly consistent with the dynamics of the originating strain of SARS-CoV-2: $\beta_s = 0.8/\mathrm{day}$, $\beta_a = 0.75 \beta_s$, $1/\nu=2\ \mathrm{days}$, $1/\gamma_s=1/\gamma_a=5\ \mathrm{days}$, and $f=0.01$ \citep{park2020reconciling}.
Under this parameterization, we have symptomatic and asymptomatic reproduction numbers of $\RR_s = 4$ and $\RR_a = 3$.

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{c|l|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Parameter} & \textbf{Description} & \textbf{Assumed values}\\
      \hline
      $\beta_s$ & Symptomatic transmission rate & See Materials and Methods\\
      \hline
      $\beta_a$ & Asymptomatic transmission rate & See Materials and Methods \\
      \hline
      $\beta_p$ & Presymptomatic transmission rate & See Materials and Methods \\
      \hline
      $1/\nu$ & Mean latent period & $2\ \mathrm{days}$\\
      \hline
      $1/\sigma$ & Mean presymptomatic infectious period & $2\ \mathrm{days}$\\
      \hline
      $1/\gamma_s$ & Mean symptomatic infectious period & $3\ \mathrm{days}$\\
      \hline
      $1/\gamma_a$ & Mean asymptomatic infectious period & $3\ \mathrm{days}$\\
      \hline
      $p$ & Proportion asymptomatic & 0--1\\
      \hline
      $f$ & Fatality rate for symptomatic case & 0.01\\
      \hline
      $\delta$ & Reduction in symptomatic transmission rate & 0--1\\
    \end{tabular}
    \caption{Paramter descriptions and values for the generalized asymptomatic model.}
    \label{tab:table1}
  \end{center}
\end{table}

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{c|l|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Parameter} & \textbf{Description} & \textbf{Assumed values}\\
      \hline
      $\beta_s$ & Symptomatic transmission rate & $0.8/\mathrm{days}$\\
      \hline
      $\beta_a$ & Asymptomatic transmission rate & $0.75 \beta_s$ \\
      \hline
      $1/\nu$ & Mean latent period & $2\ \mathrm{days}$\\
      \hline
      $1/\gamma_s$ & Mean symptomatic infectious period & $5\ \mathrm{days}$\\
      \hline
      $1/\gamma_a$ & Mean asymptomatic infectious period & $5\ \mathrm{days}$\\
      \hline
      $p$ & Proportion asymptomatic & 0--1\\
      \hline
      $f$ & Fatality rate for symptomatic case & 0.01\\
      \hline
      $\delta$ & Reduction in symptomatic transmission rate & 0--1\\
      \hline
      $\epsilon_i$ & Protection against infection & 0--0.8\\
      \hline
      $\epsilon_s$ & Protection against symptoms & 0--0.8\\
      \hline
      $\epsilon_d$ & Protection against deaths & 0--0.8\\
    \end{tabular}
    \caption{Paramter descriptions and values for the asymptomatic model with immunity.}
    \label{tab:table1}
  \end{center}
\end{table}

>>>>

% \usepackage{natbib}
\bibliographystyle{Science}
\date{\today}

@@

\begin{flushleft}{
	\Large
	\textbf\newline{
		Intermediate levels of asymptomatic transmission can lead to the highest levels of fatalities
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1},
Jonathan Dushoff\textsuperscript{2,3,4},
Bryan T. Grenfell\textsuperscript{1,5},
Joshua S. Weitz\textsuperscript{6,7,8,*}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Princeton School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\textbf{6} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{7} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{8} Institut de Biologie, \'{E}cole Normale Sup\'{e}rieure, Paris, France
\\
\bigskip

@@

Asymptomatic transmission and fatalities

@@

Asymptomatic infections improve individual-level outcomes but increase overall transmission, thereby causing more deaths.

@@

Asymptomatic infections have hampered the ability to characterize and prevent the transmission of SARS-CoV-2 throughout the pandemic.
Although asymptomatic infections reduce severity at the individual level, they can make population-level outcomes worse if asymptomatic individuals---unaware they are infected---transmit more than symptomatic individuals.
Using an epidemic model, we show that intermediate levels of asymptomatic infection lead to the highest levels of epidemic fatalities when the increase in asymptomatic transmission, due either to individual behavior or mitigation efforts, is strong. 
We generalize this result to include presymptomatic transmission, showing how intermediate levels of non-symptomatic transmission lead to the highest levels of fatalities.
Finally, we extend our framework to illustrate how the intersection of asymptomatic spread and immunity profiles determine epidemic trajectories, including population-level severity, of future variants.

@@

SARS-CoV-2 has had devastating effects at the population level.
However, many individuals experienced mild cases, making it harder to estimate the magnitude of spread and fatality rate \cite{nogrady2020data}.
The ratio of fatalities to documented cases (the case-fatality rate, CFR) is typically between 1\%--4\%, varying across population because of testing patterns, treatment practice, case definitions, and other factors \cite{rajgor2020many,VERITY2020669,yang2020early}.
But many infections are never documented;
the ratio of fatalities to total infections (the infection fatality rate, IFR) has been estimated to be closer to 0.5\%--1\% for pre-vaccinated populations whose demographics are similar to those of the United States \cite{levin2020assessing}. 
This means that more than 99\% of individuals infected with COVID-19 will survive. 
Moreover, at least half of the infections are sufficiently mild that they could be classified as subclinical or even asymptomatic.

@@

Early in the pandemic, a COVID-19 outbreak on the Diamond Princess cruise ship played a critical role in understanding the role of asymptomatic infections in the spread of SARS-CoV-2;
the outbreak occurred among 3711 passengers and crew, of whom 634 individuals tested positive by 20 February 2020 \cite{mizumoto2020estimating}.
It has been estimated that 75\% (95\% C: 70\%-78\%) of all infections on the cruise ship were asymptomatic (\fref{evidence}A) with about half of total infections undetected \cite{emery2020}.
The relative transmission rate of asymptomatic individuals aboard the Diamond Princess was not well constrained, but low relative transmission rate (below 25\%) by asymptomatic individuals was ruled out because it required unrealistically high transmissibility for symptomatic individuals (\fref{evidence}B).
Modeling studies have typically assumed that transmissibility is lower for asymptomatic than for symptomatic individuals; assumptions have ranged from 10\%--100\% \cite{ferretti2020quantifying,lavezzo2020}.
Similarities in viral load trajectories of asymptomatic and symptomatic individuals provide indirect support for the transmissibility of asymptomatic individuals (\fref{evidence}C, \cite{Kissler2020}); 
however, differences between inferred total viral load from Ct values and infectious viral load add uncertainties to how well asymptomatic individuals can transmit relative to that of symptomatic individuals \cite{romero2021ct}.
We note also that asymptomaticity is expected to be more heterogeneous in a diversity of outbreak settings \cite{alene2021magnitude}.
For example, during the early pandemic, Davies \textit{et al.}'s analyses of surveillance data across six countries revealed that older individuals were less likely to have subclinical infections (\fref{evidence}D), providing indirect evidence for heterogeneity in asymptomaticity \cite{davies2020}.
Differences in contact rates between age classes further contribute to the heterogeneity in asymptomatic transmissibility.
For now, we primarily focus on a homogeneous population and return to the age effect in discussing our model-based findings.

Despite quantitative uncertainties in asymptomatic transmissibility, individuals infected asymptomatically with SARS-CoV-2 can still transmit to others. 
This means that the presence of asymptomatic infections may have countervailing effects at the population level. 
On one hand, an asymptomatic infection means that the individual infected avoids hospitalization and death. 
On the other hand, asymptomatic infections are less likely to be detected \cite{fraser2004factors,moghadas2020implications}, meaning that asymptomatic individuals are less likely to take precautions and relatively more likely to infect others;
asymptomatic SARS-CoV-2 infections present additional challenges to managing overall disease burden due to the possibility of long COVID \cite{xie2022long}.
Altogether, the prevalence of asymptomatic infections could paradoxically make population-level outcomes worse than if SARS-CoV-2 was more dangerous at the individual level.

@@

High values of $\delta$ required for the nonlinear effects of asymptomaticity on deaths may seem unrealistic.
For this particular model, it does not make biological sense for $\delta$ to be greater than the amount of post-symptomatic transmission, because pre-symptomatic transmission is implicitly included in the $I_s$ compartment.
While several studies have estimated the proportion of pre-symptomatic transmission to be around 30\%--60\% for the SARS-CoV-2 wildtype strain, many of these were likely affected by intervention and behavioral effects, as they were conducted after SARS-CoV-2 awareness became widespread \cite{he2020temporal}.
Instead, \cite{sender2021unmitigated} recently estimated that the proportion of pre-symptomatic transmission could have been as low as 20\% (95\%CI: 6\%--32\%) during the first few weeks of the pandemic when the pandemic-awareness and intervention measures were minimal.
There are two implications of this updated estimate---first, a low proportion of pre-symptomatic transmission suggests that high $\delta$ values are feasible (although not necessarily likely) during the initial pandemic phase; and second, intermediate levels of behavioral effects ($\delta > 0$) would have been already present early in the pandemic to reduce the proportion of pre-symptomatic transmission from 80\% to as low as 40\%.

@@

We now apply our framework to understand the impact of immunity on total fatalities at the population scale by dividing the population into two groups: immunologically naive and protected.
For simplicity, we do not distinguish whether the immunity is derived from natural infections or vaccines.
The dynamics of immunologically naive individuals are equivalent to our original model (\fref{base}).
The dynamics of protected individuals include three additional parameters, which characterize the amount of protection against infection $\epsilon_i$, symptoms (given infection) $\epsilon_s$, and deaths (given symptoms) $\epsilon_d$ (\fref{immune}).
For simplicity, we assume that the population is split in half (50\% naive and 50\% protected) and mixes homogeneously. We also do not consider the separate effect of immunity on transmission (beyond the effect on infection). 
In other words, we assume that asymptomatic infections in protected and unprotected people have the same reproduction numbers (and likewise for the symptomatic infections).
In practice, both asymptomatic and symptomatic infections in protected people are less likely to transmit than their unprotected counterparts \cite{lipsitch2020understanding}: asymptomatic infections in protected people may indicate limited viral replication or even immune boosting, in which case an exposed individual may successfully fight off the pathogen early in infection before it can be transmitted; and symptomatic infections in protected people may reflect a strong immune response (rather than high viral load), in which case symptomaticity can be a poor proxy for transmission.
We assume a relatively strong behavioral effect $\delta=0.8$ for illustration.

@@

The outcomes in our simulations of invading variants resemble the dynamics of the SARS-CoV-2 Omicron variant.
Despite moderate levels of vaccine effectiveness against symptomatic and reduced levels of severe cases caused by the Omicron variant, especially after booster shots \cite{andres2022omicron}, both vaccine- and infection-derived immunity provided limited protection against infections \cite{pearson2021omicron}.
This immune evasion helped the Omicron variant to cause more infections in South Africa than previous variants \cite{sun2022omicron}.
Moreover, even though the Omicron variant is probably milder than the Delta variant \cite{MENNI20221618,ulloa2022estimates}, the number of hospitalizations and deaths caused by the Omicron variant was higher than those caused by the Delta variant in many locations \cite{Iacobuccio254,faust2022omicron,sigal2022estimating}.

There are several limitations to our analysis.
First, while we are able to generalize the model to include both pre-symptomatic and asymptomatic transmission, behavioral and intervention effects must be relatively large in order for the fatality to peak at intermediate levels of asymptomaticity (typically requiring a reduction in transmission rate of 60\% or more for most of our chosen parameter sets).
Second, the model framework is able to incorporate the impacts of immunity of infection, symptoms, and severity,
but we neglected the effects of immunity on transmission, which also has important effects on disease dynamics \cite{saad2020immune,gallagher2021indirect}.
In particular, if immunity provides stronger protection against transmission among immune individuals, population-level outcomes will be better than what our model predicts.
Estimating protection against different endpoints (e.g., infection, symptom, death, and transmission) can help narrow this uncertainty.
Finally, we assumed that asymptomatic and symptomatic individuals are infected for the same amount of time.
Analysis of viral load trajectories suggests that asymptomatic individuals may clear infections faster \cite{Kissler2020};
however, asymptomatic individuals may still transmit for a longer period of time if symptomatic individuals self-isolate quickly after symptom onset.
The individual-level differences in the asymptomatic and symptomatic transmission time scale can have important implications for the inferences and predictions of pathogen dynamics \cite{park2020time,harris2022time};
nonetheless, we expect that predictions on the final size of the epidemic and total fatalities will be robust to small differences in the transmission time scale between asymptomatic and symptomatic individuals.

Even though we assumed a homogeneous population throughout, our analysis also has important implications for age-dependent heterogeneity in asymptomaticity (as shown in \fref{evidence}D).
For example, vaccinations and intervention measures primarily targeting older individuals can prevent severe infections and improve individual-level outcomes.
However, asymptomatic individuals, especially younger individuals with high contact rates, can still transmit to other, older individuals, potentially making population-level outcomes worse than they would be if intervention measures were distributed differently.
We note that other factors, such as the efficacy of a vaccine and types of immunity provided by the vaccine, also play critical roles in making these decisions---in many cases, protecting the most vulnerable will be the optimal decision to minimize deaths \cite{moore2021modelling}.

In summary, using a series of simplified models, we have shown that asymptomatic infections (or, more generally, non-symptomatic transmission) can represent a double-edged sword leading to a better outcome for many individuals while facilitating onward transmission that leads to a worse outcome for the population as a whole.
Extending our framework further shows that immunity profile (i.e., reduction of infection, symptoms, and/or severity due to immunity) plays a critical role in determining the dynamics of future variants;
these results extend previous work on post-pandemic trajectories that focus primarily on cross-immunity \cite{kissler2020projecting,lavine2021immunological}.
For example, while protection against symptoms unaccompanied by protection against transmission protects health at the individual level, it can lead to more infections, and potentially more deaths, at the population level.
A similar concern was raised in prioritizing vaccine choices that could reduce severe outcomes vs. others that could reduce transmission \cite{koelle2022changing}.

SARS-CoV-2 has proven hard to control in large part because transmission is often decoupled from symptoms. 
% Mitigation efforts have often prioritized responding to symptoms---including symptom-based testing, fever checks, mask-wearing for infectious individuals---a different approach that strives to reduce the chance of asymptomatic transmission while increasing treatment of symptomatically infected individuals could both reduce infection risk at the source and the overall rates of severe outcomes for the population as a whole.
Our model reinforces the need for dual approaches---prioritizing the reduction of asymptomatic spread (e.g., via risk awareness campaigns \cite{Bavel2020,chande2020real,sinclair2021pairing}, asymptomatic testing programs \cite{mina2021covid,gibson2022surv,Ranoa2022}, mask-wearing indoors and in crowded environments \cite{jones2020two,prather2020reducing,howard2021ev}, and through improvements in ventilation \cite{allen2021indoor,wang2021airborne}) while improving the treatment of symptomatic cases, particularly amongst older individuals at highest risk for severe outcomes.
Given the link between age and asymptomatic infections \cite{davies2020}, interventions may consider different approaches in strongly age-structured populations (e.g., schools or long-term care facilities).
Mass vaccination is also expected to be important especially if future vaccines induce more transmission blocking.
As more variants continue to emerge, monitoring the impacts of preexisting immunity (whether through vaccination and/or infections  \cite{lopman2021framework}) on preventing infections, and not just disease, will be critical to controlling the course of the pandemic.

@@

First, we consider a compartmental model with asymptomatic and symptomatic infections in a homogeneously mixing population.
The model dynamics are as follows:
\begin{align}
\dot{S} &= -\beta_a S I_a -(1-\delta) \beta_s S I_s \\
\dot{E} &= \beta_a S I_a + (1-\delta) \beta_s S I_s - \nu E\\
\dot{I}_a &= p \nu E - \gamma_a I_a\\
\dot{I}_s &= (1-p) \nu E -\gamma_s I_s\\
\dot{R} &= \gamma_a I_a + (1-f) \gamma_s I_s \\
\dot{D} &= f \gamma_s I_s
\end{align}
where the transmission rate $\beta$ and removal rate $\gamma$ can be potentially differ between asymptomatic and symptomatic individuals.  
Here, $\delta$ denotes the reduction in transmissibility due to responsive measures taken by symptomatic individuals.
Throughout the paper, we use parameters that are broadly consistent with the dynamics of the originating strain of SARS-CoV-2: $\beta_s = 0.8/\mathrm{day}$, $\beta_a = 0.75 \beta_s$, $1/\nu=2\ \mathrm{days}$, $1/\gamma_s=1/\gamma_a=5\ \mathrm{days}$, and $f=0.01$ \cite{park2020reconciling}.
Under this parameterization, we have symptomatic and asymptomatic reproduction numbers of $\RR_s = 4$ and $\RR_a = 3$.

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{c|l|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Parameter} & \textbf{Description} & \textbf{Assumed values}\\
      \hline
      $\beta_s$ & Symptomatic transmission rate & See Materials and Methods\\
      \hline
      $\beta_a$ & Asymptomatic transmission rate & See Materials and Methods \\
      \hline
      $\beta_p$ & Presymptomatic transmission rate & See Materials and Methods \\
      \hline
      $1/\nu$ & Mean latent period & $2\ \mathrm{days}$\\
      \hline
      $1/\sigma$ & Mean presymptomatic infectious period & $2\ \mathrm{days}$\\
      \hline
      $1/\gamma_s$ & Mean symptomatic infectious period & $3\ \mathrm{days}$\\
      \hline
      $1/\gamma_a$ & Mean asymptomatic infectious period & $3\ \mathrm{days}$\\
      \hline
      $p$ & Proportion asymptomatic & 0--1\\
      \hline
      $f$ & Fatality rate for symptomatic case & 0.01\\
      \hline
      $\delta$ & Reduction in symptomatic transmission rate & 0--1\\
    \end{tabular}
    \caption{Paramter descriptions and values for the generalized asymptomatic model.}
    \label{tab:table2}
  \end{center}
\end{table}

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{c|l|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Parameter} & \textbf{Description} & \textbf{Assumed values}\\
      \hline
      $\beta_s$ & Symptomatic transmission rate & $0.8/\mathrm{days}$\\
      \hline
      $\beta_a$ & Asymptomatic transmission rate & $0.75 \beta_s$ \\
      \hline
      $1/\nu$ & Mean latent period & $2\ \mathrm{days}$\\
      \hline
      $1/\gamma_s$ & Mean symptomatic infectious period & $5\ \mathrm{days}$\\
      \hline
      $1/\gamma_a$ & Mean asymptomatic infectious period & $5\ \mathrm{days}$\\
      \hline
      $p$ & Proportion asymptomatic & 0--1\\
      \hline
      $f$ & Fatality rate for symptomatic case & 0.01\\
      \hline
      $\delta$ & Reduction in symptomatic transmission rate & 0--1\\
      \hline
      $\epsilon_i$ & Protection against infection & 0--0.8\\
      \hline
      $\epsilon_s$ & Protection against symptoms & 0--0.8\\
      \hline
      $\epsilon_d$ & Protection against deaths & 0--0.8\\
    \end{tabular}
    \caption{Paramter descriptions and values for the asymptomatic model with immunity.}
    \label{tab:table3}
  \end{center}
\end{table}
